
This body of work was carried out during my time as a Ph.D student at Okinawa Institute of Science and Technology Graduate University (OIST), and grew out of work and ideas I started to pursue at University College Cork (UCC), Ireland. The purpose of the work is to understand the dynamics of rapidly rotating Bose--Einstein condensates subjected to perturbations, and to control and engineer specific non-equilibrium states during the evolution. While it is possible to derive some analytical solutions for rapidly rotating condensates (e.g. lowest Landau level), such solutions are rare. This thesis concentrates on the numerical solutions of the Gross--Pitaevskii equation, and the resulting dynamics within this framework. It focuses on gaining an understanding of the dynamical behaviour of quantum vortices in an Abrikosov geometry subjected to perturbation. %A further application of this work would be the use of such tools to model and investigate turbulent and chaotic quantum behaviour.

Understanding ultracold Abrikosov vortex systems can help with engineering quantum states for future technologies. Ideally, they can be used for long-term memory storage in quantum computing systems as they are topologically protected and are very robust. They also allow the study of quantum mechanical effects on mesoscopic scales, and the inherent periodicity makes them a great tool for simulating condensed matter physics. Furthermore, perturbed vortex lattices can be used to investigate turbulent, and possibly chaotic, quantum behaviour. While turbulent classical systems are notoriously hard to understand and control, quantum turbulence is thought to offer a more controllable route due to the quantisation condition of the circulation. It is therefore of large interest to develop new tools for manipulating and engineering specific states or rotating condensates. In this work I concentrate on two types of perturbations to the equilibrium state: i) the modification of the phonon spectrum of the condensate which does not influence the angular momentum, in particular through the use of a kicked optical potential; ii) the direct control of the topological excitations, and hence the angular momentum. This is performed with direct phase engineering of the wavefunction. I examine both in this order, and investigate their usefulness for generating the states described.

To do this I assume a trial system of a rapidly rotating BEC having a large number of vortices, arranged in a triangular Abrikosov lattice pattern. The wavefunction is subjected to the different perturbances listed earlier, and I examine the resulting dynamics. This system requires the solution of a two-dimensional partial differential equation at high grid resolution subjected to a variety of different initial conditions and controllable perturbations. This is a non-trivial numerical problem, and requires the use of advanced numerical computing techniques to allow for results in a reasonable time. For this I make use of graphics processing unit (GPU) computing, and I will discuss the development of such tools and compare them against conventional simulation techniques.

The thesis is organised as follows:

\section{Background}
I will first give a brief introduction to the field of cold-atomic gases, and discuss the theoretical framework to describe Bose--Einstein condensation. Emphasis will be placed on material and works relevant to the studies I have performed in this thesis. I will present a derivation of the Gross--Pitaevskii equation, used to model Bose--Einstein condensates, as well as a discussion of the Bogoliubov-de Gennes equations. I will then discuss the hydrodynamic description of the condensate, and give the hydrodynamic form of the Gross--Pitaevskii equation. Here I introduce superfluidity, and the nature of quantised vortices in these systems. I will conclude with an outlook on the cutting edge work in the field in the context of condensate trapping and control.


\section{Numerical methods}
In this chapter I will discuss methods for numerically solving the Gross--Pitaevskii equation for simulating the dynamics of Bose--Einstein condensates. The Fourier split-operator method will be introduced, as well as the need for imaginary time evolution, and considerations required to effectively simulate the condensate. Graphics processing unit (GPU) computing will be introduced here, with the implementation of the Gross--Pitaevskii equation discussed. To demonstrate the power of GPU computing I will present a difficult numerical problem here, namely the solution of an experimentally realistic situation for a single, ultracold atom on an atomchip, for which the treatment of the fully three-dimensional Schr\"odinger equation is required. The use of GPU computing makes this problem tractable in realistic times. The work focuses on the area of adiabatic control techniques, and demonstrates the use of GPU computing to describe the long-time dynamics of a system for observing matter-wave spatial adiabatic passage. Relevant works to this discussion have been published in \cite{AO:Morgan_pra_2013}, and source available at \cite{MLXD_GPUE}.

\section{Bose--Einstein condensate dynamics}
In this chapter I will examine the dynamics of Bose--Einstein Condensates under rotation, and discuss the methods used for probing the condensate system. I begin by introducing the dynamical behaviour of the condensate in the presence of vortices, and introduce a model system used for further discussions. I present the velocity profiles and discuss some of the behaviour expected in a condensate with many vortices. This will be followed by an introduction of the two main perturbation methods for the condensate that I will later use: optical kicking, and phase imprinting. I will also discuss the analysis of the condensate dynamics, concentrating primarily on the kinetic energy spectrum, and Delaunay triangulation and Voronoi tesselation.

\section{Moir\'e superlattice structures}
Here I investigate effects stemming from the optical kicking of a condensate carrying a vortex lattice. The robustness of the vortex lattice after a kick with an optical potential of the same geometry as the vortex lattice is demonstrated by showing little to no deviation of ideal vortex positions in the dynamics. The resulting condensate density however shows the appearance of a superlattice pattern. I analyse this system, and demonstrate that the resulting superlattice pattern stems from interference between the optical kicking potential and the present vortex lattice in reciprocal space. Moir\'e interference theory accurately predicts the observed behaviour, and is backed up by examining the kinetic energy spectrum of the condensate. Uses of this technique are discussed to conclude. The results presented in this chapter have been published in \cite{VTX:oriordan_pra_2016}.

\section{Defect engineering of the vortex lattice}
To investigate the robustness of a vortex lattice in a rapidly rotating BEC, I will in this chapter discuss the effect of perturbations induced by adding or removing angular momentum through phase imprinting. Through phase imprinting, lattice vacancies and topological defects are created in the vortex lattice, and their behaviour investigated over long times. I show that the vortex lattice demonstrates highly robust behaviour, even in the presence of such defects. I discuss the use of this method for creating varying degrees of disorder in the lattice, and propose it as a system for investigating order/disorder transitions. The results in this chapter are currently submitted to Physical Review A for publication.

\section{Conclusions \& outlook}
In this chapter I conclude the work discussed in the thesis, and discuss extensions, and future ideas for the field.
