
\chapter{Guidelines on the preparation of theses} \label{ch-1}

These guidelines set out the organization and formatting requirements of the OIST Ph.D. thesis, in order to assist students in the preparation of theses for submission. The academic requirements of the thesis are defined in full in the document ``PhD Completion Requirements''. 

This particular documents refers specifically a thesis written in \LaTeX. As such, some points from the full guideline (for example page sizes) are not referenced directly here as they are already defined in this template. Some other points concerning specific pages (for example the abstract) are described in the specific pages themselves in this PDF.

\section{Guidelines on the preparation of theses}

\textbf{Title Page}:  This page is the first printed page, and can be duplicated in whole or part as the front cover in the final bound version.

\textbf{Choice of Title}: Select a descriptive and unique title that clearly communicates your research.  Avoid brief or misleading titles.  The title will be displayed on your graduation testamur.  The title should be unique within OIST, to distinguish your thesis from those of others working on similar topics.

\textbf{Declaration of Original Authorship}:  You must declare that the work is your own, and original,  by signing and including the declaration.  

\textbf{Plagiarism and Fraud}:  Students are reminded that they must take all necessary precautions to avoid plagiarism and fraudulent misrepresentations of data. 

\textbf{Co-authorship}:  Co-authors are not permitted on an OIST PhD thesis. All research and analysis should be your own work. Where co-authors have contributed to papers arising from the work, you should not include their data unless it is essential for the scientific narrative.  In such cases, full disclosure of the contribution is required. Acknowledge any work performed by others, whether at OIST or outside OIST.

\textbf{Main body}:  The main body may be arranged as a single body of material, divided into sub-sections of Introduction (including a statement of the problem), Methods, Results, Discussion, or if preferred, in chapters that each deal with a smaller part of the research, each one itself divided into sub-chapters of Introduction, Methods, Results, Discussion (or similar), as appropriate. 

\textbf{Reference List}:  Provide a complete numbered list of all material cited in your thesis using BibTeX or BibLaTeX.

\textbf{Appendices (published papers)}: The printed versions of the thesis may include, as an appendix, published papers; unpublished manuscripts that have been submitted for publication; and manuscripts ready for, or very close to, journal submission.  These should be placed immediately after the final pages of the thesis, and separated from the thesis itself by a single dividing page with the text: ``Previously published articles associated with the research described in this thesis'', or similar. These published papers are included solely for completeness of the bound thesis as a historical record of your achievement, and are for future reference of yourself and the supervisors.  They will not be included in the on-line version of the thesis. Papers co-authored during the period of the thesis that do not include material presented in this work should not be included.

\section{Formatting Requirements}

\textbf{Equations}: Equations are considered to be part of the text; they should be formatted consistently throughout the thesis, following the advice of the student?s supervisor.  

\textbf{Spelling}: American spelling should be used.

\textbf{Printing}:  Theses should be printed on one side of a page only (so that when bound, the right hand page is the printed page).  This avoids text and image bleed through.  Temporary bindings may use any reasonable white bond paper, in A4 size.  Laser printing (in black ink wherever possible, and colour for images where necessary), should be used exclusively, rather than alternatives such as ink-jet, dye sublimat ion, or wax transfer (for durability of the print).  Bound copies should use acid free paper (also known as archival paper) to ensure longevity of the thesis in the collection.

\textbf{Colors}:  Colors may be used in images and charts where necessary to enhance comprehension, but should not be used for normal text or headings.  Avoid the exclusive combination of red and green for binary images, to assist those who have difficulty discriminating hues.  All text should be in black unless color-coding is necessary for meaning or contrast.

\textbf{Word length}:  No minimum word length is imposed on OIST graduate theses.  However, be concise in your language and succinct in your expression.  The average length of a PhD thesis will vary between fields and between authors, but typical PhD theses are 200-400 pages in length (approximately 40-80,000 words of main body text).  

\textbf{Citations}:  All papers that you reference in your work must be referenced in full using a style relevant to your field. Provide the full title, a complete list of authors, and the article location and year of publication in the same style for all papers. Use one of several styles you have been introduced to in previous writing.  Refer to papers in the text by either a numerical superscript, a bracketed number or by reference to (Author et al., 1999).  Be consistent in your citation style throughout all sections of your thesis.  Provide a complete list of all papers, books, and proceedings cited in your thesis at the end of the main body of text.  Do not include papers in this list that were not cited in the thesis.  Reference manager software such as Endnote or similar programs that offer ``Cite-While You Write'' functionality can assist this process greatly if you are using Word. Use BibTeX or BibLaTeX if you are using LaTeX.

Citing one reference can be done like so: \cite{Lee98} and multiple references in one go like so \cite{Fil09, Muc10, Kra27}.

\textbf{Editing}:  The thesis should be entirely your own work. Minimal editing may be provided by your supervisor(s) or peers but only as a review of initial drafts.  Do not seek assistance from OIST internal or paid external editing services, unless directed to do so by the Dean in revision stages.

\section{Intellectual property and copyright}

The student will retain copyright of the published work, in perpetuity.  The student acknowledges that OIST remains the owner of the intellectual property generated by the research presented in the thesis and that publication of the thesis under the author?s copyright does not diminish this claim.  The thesis will be published online in electronic form within one year of graduation.


 \section{Submission of temporary and bound versions}

The thesis will be submitted first in printed form in a style formatted to assist the examination process (single side printing, double spacing, temporary binding).  Two copies of the examination version must be bound in temporary form (using the comb binder available in the graduate school office, or with a glued spine cloth).  Soft plastic end covers may be used. This version should include (inserted after the thesis itself)  any papers published during the tenure of the thesis that are relevant to the material therein, for the convenience of the examiners.  A digital version of the same file (without the ``published papers'' appendix) should also be sent to the Grad School with your submitted version, which will be kept as a backup.

The two external examiners will examine the written thesis and prepare a report.  This report may ask for minor revisions, but it is not expected that major revisions will be undertaken at this point. When the two copies of the marked thesis are returned (with annotations) the student will be asked to make any required minor revisions, and must then confirm with the Graduate School that the revisions are accepted and complete within one month.  Once the final version of the thesis is approved, the student will then be eligible to defend the thesis.  No additional changes may be made at that point.  The final version of the thesis must be submitted as an electronic file in PDF format.  Once the thesis defense has been concluded, and the degree awarded, the electronic file will be published online by the graduate school in the OIST Institutional repository.

The student should also prepare a permanent bound version of the final approved thesis (essentially the same as the PDF with altered margins) to present to the supervisor(s), according to the format described below (appendix 1, 2, and 3).  The bound version must be submitted before the degree may be awarded, and may be submitted either before or after the defense.  The graduate school will cover necessary costs for binding the supervisor copy (or copies).
