\unnumberedchapter{Acknowledgment}
\chapter*{Acknowledgment}

Theses must acknowledge assistance received in any of the following areas:

\begin{itemize}
\item Designing the research
\item Executing the research
\item Analyzing the data
\item Interpreting the data/research
\item Writing, proofing, or copyediting the manuscript
\end{itemize}

Where co-authors have contributed to papers arising from the work, you should not include their data unless it is essential for the scientific narrative.  In such cases, full disclosure of the contribution is required. Acknowledge any work performed by others, whether at OIST or outside OIST.


\iffalse
Firstly, I would like to express my sincerest gratitude to my mother, nan, and Maria who have always been there for me, and provided great support from across the planet since I started my PhD. Next, my sincerest gratitude goes to Prof. Thomas Busch, for freeing me from the shackles of a boring existence and introducing me to the (ultra)cool world of atomic physics ... and essentially dragging me to Japan for a few years. To past and present members of both the Ultracold Quantum Gases group at University College Cork, and Quantum Systems Unit at OIST, thanks for keeping me sane (most of the time). In particular, thanks to Tadhg, despite having skipped out on Quantum Tokyo Team 2012, for letting me fix your code and formatting choices again and again; Mossy, for your tender love of Nic Cage, and singing cheesy 80's songs; Steve, for never drinking alcohol; you are my inspiration; Tara, for ...

Angela for always putting up with my stupid questions, helping out when I've been thoroughly confused, and also, for baking awesome cakes.
\fi
